WHS Machine Learning: Lessons Learned

### 2013-06-21 paper/paper.tex
End of the year commit: paper and demos complete

\documentclass[11pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{csvsimple}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage[table]{xcolor}

@@

\author{Kevin Gao, Matt Li, Ashvin Nair, Saavan Patel}
\title{WHS Machine Learning: Lessons Learned}

@@

Introductory machine learning is the subject of a yearlong autodidactic pioneering computer-science course at Winchester High School. The intent of this course is to understand the algorithms of machine learning and implement them in code. In this course we studied decision trees, genetic algorithms, hidden Markov models, and artificial neural networks. Whenever possible, we applied the algorithms to real-world data.

The purpose of this paper is to publicize simple implementations of introductory machine learning algorithms in an accessible and clear way. For additional reference, our implementation of the below algorithms can be found at \url{https://github.com/anair13/whs-machine-learning}.

@@

Decision trees use data to make predictions. Their purpose is to produce an answer while asking as few questions as possible, which is done by figuring out the most informative question at every point. This quickly narrows down the information into a likely answer.

@@

For reference, we create a set of variables called attributes consisting of a name and possible values represented by the range from 0 to (number of values - 1). These attributes act as questions to ask about. For example, an attribute named ``temperature'' may have the values ``hot,'' ``mild,'' and ``cool'' or $\{0, 1, 2\}$. Most of the attributes are determining factors, but the very last attribute is the outcome, such as ``play tennis'' with values of ``yes'' or ``no.'' The value must be given when training the model, but it can be predicted from the other attributes once the model is constructed. The data itself is a matrix with each row being a separate case and each column containing values for an attribute. We get the information from a text file of the same format, gather all the values that appear for each attribute, assign them numbers, and finally parse the data into a two-dimensional array.

@@

\[
H(X) = \sum{(-p^+ \log_2{p^+} - p^- \log_2{p^-})}
\]

where $p^+$ is the proportion of positive examples, and $p^-$ is the proportion of negative examples in the set.

@@

where the sum iterates through every set of positive examples $V$ (the different sets the question splits the data into). The ``best question'' is the attribute with the highest information gain.

@@

To create the tree, we use Node objects to represent each node of the tree. The objects contain one attribute and have children nodes. We use the algorithm with the original data to find the first question and place the attribute at the root node. For each of the attribute's values, we comb through the data and find only the cases with the same value for the attribute and create a new data set with the attribute removed. The algorithm is used again on the new data to figure out the next question. This attribute is then put into the child node for that value. The tree is grown recursively until we can predict the outcome with 95\% certainty.

@@

On the right is the data. After determining that outlook is the best question to ask first, we group the data based on the value for outlook. Since we do not want to ask the same question again, the data for outlook is removed. The red cells become the data used to decide the question for the node at the ``sunny'' branch, green for the ``overcast'' branch, and blue for the ``rainy'' branch.

\newpage
\subsection{Results}
\label{DecisionTrees:Results}
}

We applied the decision tree program to car buyers, the census, congressional voting records, and CPU performance data. It was usually able to give a confident answer within just a handful of questions out of a dozen or more attributes. This varied greatly with the strength of association between attributes and whether the specific case followed the trends of others.

@@

We used Weka to check our results, and our trees were almost identical to theirs, with small differences in the level of pruning based on the threshold we used for certainty.

@@

Decision trees are an efficient way to crunch a lot of information down to a few key questions that can fairly accurately predict an outcome. While it becomes harder to use when the number of attributes becomes extremely large, its advantage over many other predicting algorithms is that it requires less instances of data to create a useful model. The simple structure of the decision tree makes it a great introduction to machine learning.

\section{Genetic Algorithms}
\label{Genetic Algorithms}

Genetic algorithms find parameters to optimize a given function. We used it in combination with the Hungarian algorithm to create a program that selects an optimal math team given individual mathlete scores. In this section, we explain our implementation of a genetic algorithm, as well as explain our larger program.

The genetic algorithm is a computational implementation of the theory of natural selection in nature. The idea behind the algorithm is ``survival of the fittest,'' where the fittest individuals are the most likely to survive to the next generation. The idea of mutations is also introduced, where random mutations occur within the population that introduce variability, and reduce the chances of the population being stuck in a local maxima, when global maxima are available.

\begin{figure}[h]
  \begin{center}
	\includegraphics[scale=0.6]{natural_selection.png}
  \end{center}
  \caption{A simple description of the theory behind natural selection}
\end{figure}

\subsection{Data}
\label{Genetic Algorithms:Data}

Genetic algorithms attempt gradient ascent of parameters by mimicking natural selection. Given a fitness function to be optimized, an individual is defined as a set of parameters for the function and we call each parameter an ``attribute" of that individual.

@@

Generally, a genetic algorithm optimizes a function with a large set of parameters when the result of a parameter change is unknown. To start a genetic algorithm, a population,$N$, of random individuals with random traits is generated. Each individual represents a set of parameters to be optimized. The goal is to generate a population of almost homogeneous individuals that all represent the fittest possible member of the population.

Fitness (the fitness of individual $I$ will be denoted $F_{I}$) is a function that measures the success of the individual: higher fitness means that the individual is more likely to reproduce.

In addition, the algorithm depends on:
\begin{itemize}
\item Crossover rate $C$, the probability a trait is crossed over during breeding
\item Mutation rate $M$, the probability that a trait is randomly changed. Mutations are used to prevent individuals from converging at local extrema and therefore miss global extrema.
\item Breed proportion $B$, the proportion of each population that is generated from ``children'' of individuals in the previous population
\end{itemize}

In all, the total process runs:

\begin{enumerate}
\item An individual is selected based on its fitness, the probability is:

@@

\item The individual is either set to breed or put directly into the next generation, depending on $B$.
\item If the individual is set to breed, another individual is selected with probability $P(I)$. ``Breeding'' occurs when two new individuals are created by switching random traits between the two individuals. $C$ is the probability one of the traits is switched between individuals. In this way, two new individuals become part of the new population.
\item Individuals are selected from the previous population with probability $P(I)$ and put directly into the next generation
\item Before individuals enter the next generation, traits may be mutated with a probability $M$. 
\end{enumerate}

@@

Data was presented in the form of a csv file containing the individual scores of the 20 math team members for a try out. The goal was to choose 10 members out of the 20, each member competing in 3 out of the 5 rounds available. The csv file contains the score for each of the math team members and the score for each of the 5 rounds.

Genetic algorithm for math team selection was combined with the Hungarian algorithm to efficiently produce a result. The genetic algorithm produced a team, and the Hungarian algorithm chose which members of the team would do which rounds, which in turn went to the fitness algorithm, which calculated the fitness of the given team (the total score of the team).

@@

We implemented the Hungarian algorithm as a part of the fitness algorithm in math team selection. The Hungarian algorithm solved the optimization problem in polynomial time, which makes the rest of the genetic algorithm more efficient.

@@

A comma-separated-values file of the scores of approximately 20 math team members is entered into the algorithm, and optimal math teams are outputted through the command line. The algorithm works to pull a team of 10 math team members, and pick 3 out of the 5 available rounds for each math team member. In many cases, there are multiple optimal math teams. The algorithm is able to optimize for the math team problem within 100 generations, which takes approximately 1 minute to complete. This shows that the genetic algorithm in combination with the hungarian algorithm is effective in solving the math team allocation problem in an acceptable period of time, and could actually be useful in helping a real math team.

@@

The genetic algorithm is a useful optimization tool when the effect of parameters on the overall fitness of a function is unknown. We learned the simplest optimization tool in the machine learning environment, and provided a base for further endeavors into the field. The genetic algorithm provides a useful base for more complex and efficient algorithms, such as simulated annealing. By combining the genetic algorithm with the Hungarian algorithm to solve a real problem our math team had, we found a great use for our programming and a useful way to combine otherwise unrelated algorithms.

@@

Hidden Markov models care about sequences of events. They can be applied to data like sentences with each word tagged with a part of speech, where the innate cause of the sequence is not known but the emission of nouns and verbs is easy to identify. In our testing, we generated sequences from models, then tested whether we could predict the underlying states of those sequences and generate precise models back from those sequences.

@@

The problems are related but each method is separate. Probabilistic quantities overlap between the problems, but each problem can be neatly divided into its own solution.

@@

To find the probability of a specific observation sequence ocurring, given the model, we recursively calculate the probability of each element of the observation sequence occurring. Define the ``forward'' function $\alpha(n, i ,O|\lambda)$ as the probability of the first n observations ocurring and $Q_n$ being $S_i$.

@@

\[
C_p(q, 1) = C_p(q, 2) \pi_{q_1} B_{{q_1}{O_1}}
\]

@@

The best state sequence for $O$ is $C(\{\}, N)$.

@@

Although there are many optimizations to this process, the simplest method to extract a model from observations is counting:
\begin{itemize}
\item $\pi_i$ is the count of the first state being $S_i$ divided by the total observations
\item $a_{ij}$ is the count of transitions from $S_i$ to $S_j$ divided by the occurences of $S_i$
\item $b_{ij}$ is the count of emissions from $S_i$ to $O_j$ divided by the occurences of $S_i$
\end{itemize}

To optimize, the first quantity we must define is the probability that $Q_n$ being $S_i$. That is, $P(Q_n = S_i | O, \lambda)$. To help us, we add to our tools the backward analog of the forward algorithm. Arbitrarily, we define

\[
\beta(N, i, O, \lambda) = 1 \\
\]

Then by a similar logic as the forward algorithm:

@@

The probability of seeing the first $n$ observations and ending in $S_i$ is the forward term, the probability of seeing the rest of the observations and starting with $S_i$ is the backward term, and the observation is given so we divide by $P(O|\lambda)$. Using this function, we can incrementally improve the parameters of the model by gradient ascent.

@@

Artificial neural networks are trained through the backwards propagation of errors. The goal of backwards propagation is to adjust the weights of the connections until the error between training example output and network output is at a minimum.

@@

Artificial neural networks attempt to approximate functions. That is, using a series of training examples, a list of inputs concatenated with corresponding outputs, the network adjust weights to resemble a probable function. Inputs and outputs can easily be mapped to real world quantities. For example, we represented pictures of letters as one hundred inputs.

@@

\[
O_k(x, w) = \cfrac{1}{1 + e^{A(x, w)}} 
\]
where 
\[
A_k(x, w) = \sum_{i=0}^n{x_i*w_{ik}}
\]
$x$ is a list of inputs and $w$ is a matrix of connection weights.

@@

We chose to use the sigmoid function -- $f(x) = \cfrac{1}{1 + e^x}$ -- as a continuous alternative to a simple step function, though the step function is also commonly used. $f$ is asymptotically close to 0 for large negative values, close to 1 for large positive values, and $f(0) = 0.5$.

@@

\begin{wraptable}{l}{.45\textwidth}
\csvautotabular{pr_lows.csv}
\vspace{-120px}
\end{wraptable}

@@

Repeatedly running each training example through the network will increase the precision of the network. In order to obtain reasonable results with even simple functions, we had to run the entire set of data through the network 30 times. Below, we have our success rate plotted with additional learning cycles of the data for our character recognition project.

@@

@@

\begin{wraptable}{r}{.45\textwidth}
\vspace{-20px}
\csvautotabular{pr_caps.csv}
\csvautotabular{pr_nums.csv}
\caption{Precision, recall, and F-measure for each character identified}
\vspace{-70px}
\end{wraptable}

@@

From \citeA{web:chars-74k-hand-drawn} we found roughly 2000 sample images, 55 of each character, for our training examples. (These samples are avilable for study given that we cite \citeA{paper:deCampos09} as the original research.) In practice, we use 45 of each character to train and save 10 to test on. Using PIL, we crop each image into a square tightly containing the nonwhite regions of the image, then turn the images into 10x10 grayscale images. Each image is then represented as a list of 100 numbers: each pixel is converted into a number where white is -1 and black is 1. The output is a list of 62 numbers where each number represents a different character. The number corresponding to the correct character is 1 and the rest are 0. After training, the highest output corresponds to the letter predicted by the network.

@@

First, we tested our implementation with simple functions: given a number, return the number, its negative, and half of the number. As we ironed out the implementation, the neural net was extremely accurate in approximating these functions. We could define a neural network with our own parameters as a function to be approximated, and the training network would exactly replicate the given network within 1\% of each parameter given enough cycles through the data (In our case, 1000 cycles of 6 training examples through a network with 2 input, 2 hidden, and 2 output layers.)

@@

We found that neural networks are extremely efficient at sorting through large amounts of complicated data and finding the underlying function that we care about. However, they require a lot of data to be effective. In our case, increasing resolution of the image, and therefore increasing inputs, exponentially increases the amount of training time. On the whole, our lasting impression of neural networks is that they are an amazing tool to model convoluted functions given that we have a lot of data about the function.

\section{Conclusion}
\label{Conclusion}

The big picture we take away from this class is that by cleverly applying data to adjust parameters, machines can be trained to be really, really smart. We were taken aback by the complicated behavior: prediction, optimization, and character recognition, produced by a few simple lines of code combined with a large amount of data. Through our survey, we discovered that this field of applied statistics and computer science called machine learning is incredibly interesting. Besides the specific algorithms, we improved our implementation skills and learned new ways to think about handling and using data. We would highly recommend future classes with similarly free structure and topics.

\citeA{DBLP:journals/corr/cs-NE-0308031} and \citeA{book:pattern-classification} should force bibliography references.

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{csvsimple}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage[table]{xcolor}

@@

\author{Kevin Gao, \texttt{kmgao2@illinois.edu} \\
	Matthew Li, \texttt{matthew.yj.li@gmail.com} \\
	Ashvin Nair, \texttt{anair17@berkeley.edu} \\
	Saavan Patel, \texttt{saavan@berkeley.edu}}
\title{WHS Machine Learning: Lessons Learned}

@@

\section*{Introduction}
\label{Introduction}
Introductory machine learning is the subject of a yearlong autodidactic pioneering computer-science course at Winchester High School. The intent of this course is for students to understand the algorithms of machine learning and implement them in code. We studied decision trees, genetic algorithms, hidden Markov models, and artificial neural networks, applying the algorithms to real-world data when possible. As general reference, we used \citeA{book:machine-learning}.

The purpose of this paper is to publicize simple implementations of four different introductory machine learning algorithms in an accessible and clear way. For additional reference, our implementation of the algorithms below can be found at \url{https://github.com/anair13/whs-machine-learning}.

@@

Decision trees use data to make predictions. Their purpose is to produce an answer while asking as few questions as possible. This is done by figuring out the most informative question at every point to quickly narrow down the information to a likely answer.

@@

For reference, we create a set of variables called attributes consisting of a name and possible values represented by the range from 0 to (number of values - 1). These attributes act as questions to ask about. For example, an attribute named ``temperature'' may have the values ``hot,'' ``mild,'' and ``cool'' or $[0, 1, 2]$. Most of the attributes are determining factors, but the very last attribute is the outcome, such as ``play tennis'' with values of ``yes'' and ``no.'' The value must be given when training the model, but it can be predicted from the other attributes once the model is constructed. The data are represented as a matrix with each row being a separate case and each column containing values for an attribute. We get the information from a text file of the same format, gather all the values that appear for each attribute, assign them numbers, and finally parse the data into a two-dimensional array.

@@

\[
H(X) = -p^+ \log_2{p^+} - p^- \log_2{p^-}
\]

where $p^+$ is the proportion of positive examples, and $p^-$ is the proportion of negative examples in the set. Visually, $H = 0$ when $p^+ = 0$ or $1$ and is maximized at $p^+ = 0.5$.

@@

where the sum iterates through every set of examples with the same value for the question. The ``best question'' is the attribute with the highest information gain.

@@

To create the tree, we use node objects containing one attribute and child nodes. We use the algorithm with the original data to find the first question and place the attribute at the root node. For each of the attribute's values, we comb through the data and find only the cases with the same value for the attribute and create a new data set with the attribute removed. The algorithm is used again on the new data to figure out the next question. This attribute is then put into the child node for that value. The tree is grown recursively until we can predict the outcome with 95\% certainty.

@@

On the right are the data. After determining that outlook is the best question to ask first, we group the data based on the value for outlook. Since we do not want to ask the same question again, the data for outlook are removed. The red cells become the data used to decide the question for the node at the ``sunny'' branch, green for the ``overcast'' branch, and blue for the ``rainy'' branch.

\newpage
\subsection{Results}
\label{DecisionTrees:Results}
}
We applied the decision tree program to car buyers, the census, congressional voting records, and CPU performance data. It was usually able to give a confident answer within just a handful of questions out of a dozen or more attributes. This varied greatly with the strength of association between attributes and whether the specific case followed the trends of others.

@@

@@

Decision trees are an efficient way to crunch a lot of information down to a few key questions that can fairly accurately predict an outcome. Its advantage over many other predicting algorithms is that it requires less instances of data to create a useful model. The simple structure of the decision tree makes it a great introduction to machine learning.

\section{Genetic Algorithms}
\label{Genetic Algorithms}
{
\begin{wrapfigure}{r}{0.5\textwidth}
  \begin{center}
	\includegraphics[scale=0.6]{natural_selection.png}
  \end{center}
  \caption{A simple description of the theory behind natural selection}
\end{wrapfigure}

Genetic algorithms find parameters to optimize a given function. We use it in combination with the Hungarian algorithm to create a program that selects an optimal math team given individual mathlete scores. In this section, we explain our implementation of a genetic algorithm, as well as explain our larger program.

The genetic algorithm is a computational implementation of the theory of natural selection. The idea behind the algorithm is ``survival of the fittest,'' where the fittest individuals are the most likely to survive to the next generation. The idea of mutations is also introduced, where random mutations occur within the population that introduce variability, and reduce the chances of the population being stuck in a local maxima, when global maxima are available.

}
\subsection{Data}
\label{Genetic Algorithms:Data}

Data in Genetic Algorithms is presented in the form of a set of ``attributes'' to optimize. Each of these attributes can be changed slightly, which is how individuals are created. An individual is defined as a set of parameters of unique attributes. A fitness function is also passed in, which numerically measures how well the individual performed at the given task given that individual's parameters.

@@

The generalized Genetic Algorithm optimizes a function with a large set of parameters when the result of a parameter change is unknown.

Important terms and items in Genetic Algorithms:
\begin{itemize}
\item $F_{I}$ The Fitness of individual I, defined as the success the individual has at the given task
\item $C$ The Crossover rate, defined as the probability a trait is crossed over during breeding
\item $M$ The Mutation rate, defined as the probability that a trait is randomly changed, used to prevent local maxima from being treated as global maxima
\item $B$ The Breed proportion, defined as the proportion of each population that is generated from ``children'' of individuals in the previous population
\end{itemize}

To start a genetic algorithm, a population,$N$, of random individuals with random traits is generated. Each individual represents a set of parameters to be optimized. The goal is to generate a population of almost homogeneous individuals that all represent the fittest possible member of the population.

\begin{enumerate}
\item An individual is selected based on its fitness, the probability is simply:

@@

\item The individual is either set to breed or put directly into the next generation, depending on $B$.
\item If the individual is set to breed, another individual is selected with probability $P(I)$. ``Breeding'' occurs when two new individuals are created by switching random traits between the two individuals. $C$ is the probability one of the traits is switched between individuals.In this way, two new individuals become part of the new population.
\item Individuals are selected from the previous population with probability $P(I)$ and put directly into the next generation
\item Before individuals enter the next generation, traits may be mutated with a probability $M$. 
\end{enumerate}

@@

Data is presented in the form of a csv file containing the individual scores of the 20 math team members for a try out. The goal is to choose 10 members out of the 20, each member competing in 3 out of the 5 rounds available. The csv file contains the score for each of the math team members and the score for each of the 5 rounds.

We combine the Genetic algorithm for math team selection with the Hungarian algorithm to efficiently produce a result. The genetic algorithm produces a team, and the Hungarian algorithm chooses which members of the team would do which rounds, which in turn goes to the fitness algorithm, which calculates the fitness of the given team (the total score of the team).

@@

We implemented the Hungarian algorithm as a part of the fitness algorithm in math team selection. The Hungarian algorithm solves the optimization problem in polynomial time, which makes the rest of the genetic algorithm more efficient.

@@

A comma-separated-values file of the scores of approximately 20 math team members is entered into the algorithm, and optimal math teams are outputted through the command line. The algorithm works to pull a team of 10 math team members, and pick 3 out of the 5 available rounds for each math team member. In many cases, there are multiple optimal math teams. The algorithm is able to optimize for the math team problem within 100 generations, which takes approximately 1 minute to complete. This shows that the genetic algorithm in combination with the Hungarian algorithm is effective in solving the math team allocation problem in an acceptable period of time, and could actually be useful in helping a real math team.

@@

The genetic algorithm is a useful optimization tool when the effect of parameters on the overall fitness of a function is unknown. We learned the simplest optimization tool in the machine learning environment, and provided a base for further endeavours into the field. The genetic algorithm provides a useful base for more complex and efficient algorithms, such as simulated annealing. By combining the genetic algorithm with the Hungarian algorithm to solve a real problem our math team had, we found a great use for our programming and a useful way to combine otherwise unrelated algorithms.

@@

Hidden Markov models allow us to infer about sequences of events. They can be applied to data like sentences with each word tagged with a part of speech, where the innate cause of the sequence is not known but the emission of nouns and verbs is easy to identify. In our testing, we generated sequences from models, then tested whether we could predict the underlying states of those sequences and generate precise models back from those sequences.

@@

The problems are related but each method is separate. Probabilistic quantities overlap between the problems, but each problem can be neatly divided into its own solution. For reference, we used \citeA{paper:cornellhmm} and \citeA{book:pattern-classification} as introductions to hidden Markov models.

@@

To find the probability of a specific observation sequence occurring, given the model, we recursively calculate the probability of each element of the observation sequence occurring. Define the ``forward'' function $\alpha(n, i ,O|\lambda)$ as the probability of the first n observations occurring and $Q_n$ being $S_i$.

@@

\[
C_p(q, 1) = \pi_{q_1} B_{{q_1}{O_1}}
\]

@@

The best state sequence for $O$ is $C(\{\}, N)$. However, we further optimize the algorithm by saving $C$ for every state and time.

@@

If the states are known along with observations, the simplest method to extract a model from observations is counting:
\begin{itemize}
\item $\pi_i$ is the count of the first state being $S_i$ divided by the total observations
\item $a_{ij}$ is the count of transitions from $S_i$ to $S_j$ divided by the occurences of $S_i$
\item $b_{ij}$ is the count of emissions from $S_i$ to $O_j$ divided by the occurences of $S_i$
\end{itemize}

If states are not known, an initial model with random parameters is chosen and reestimated repeatedly. To optimize, the first quantity we must define is the probability that $Q_n$ being $S_i$. That is, $P(Q_n = S_i | O, \lambda)$. To help us, we add to our tools the backward analog of the forward algorithm. Analagous to the forward algorithm, the backward function gives us the probability of $O_{n+1..N}$ given the model and state $Q_n$. Because of the definition of the backward function,

\[
\beta(N, i, O, \lambda) = 1
\]

Then, by a similar logic as the forward algorithm:

@@

The probability of seeing the first $n$ observations and ending in $S_i$ is the forward term, the probability of seeing the rest of the observations and starting with $S_i$ is the backward term, and the observation is given so we divide by $P(O|\lambda)$. Using this function, we can incrementally improve the parameters of the model by reestimating each matrix.

@@

Artificial neural networks are trained through the backwards propagation of errors. The goal of backwards propagation is to adjust the weights of the connections until the error between training example output and network output is at a minimum. \citeA{DBLP:journals/corr/cs-NE-0308031}

@@

Artificial neural networks attempt to approximate functions. That is, using a series of training examples, a list of inputs concatenated with corresponding outputs, the network adjust weights to resemble a probable function. Inputs and outputs can easily be mapped to real world quantities. For example, we represent pictures of letters as one hundred inputs.

@@

\[
O_k(x, w) = \cfrac{1}{1 + e^{-A(x, w)}} 
\]
where 
\[
A_k(x, w) = \sum_{i=0}^n{x_i*w_{ik}}
\]
$x$ is a list of inputs and $w$ is a matrix of connection weights.

@@

We chose to use the sigmoid function -- $f(x) = \cfrac{1}{1 + e^{-x}}$ -- as a continuous alternative to a simple step function, though the step function is also commonly used. $f$ is asymptotically close to 0 for large negative values, close to 1 for large positive values, and $f(0) = 0.5$.

@@

@@

Repeatedly running each training example through the network will increase the precision of the network. In order to obtain reasonable results with even simple functions, we run the entire set of data through the network 30 times. Below, we our success rate is plotted with additional learning cycles of the data for our character recognition project.

@@

\begin{wraptable}{r}{.45\textwidth}
\vspace{-15px}
\csvautotabular{pr_caps.csv}
\csvautotabular{pr_nums.csv}
\caption{Precision, recall, and F-measure for each character identified}
\vspace{0px}
\end{wraptable}

@@

@@

From \citeA{web:chars-74k-hand-drawn} we found roughly 2000 sample images, 55 of each character, for our training examples. (These samples are available for study given that we cite \citeA{paper:deCampos09} as the original research.) In practice, we use 45 of each character to train and save 10 to test on. Using PIL, we crop each image into a square tightly containing the nonwhite regions of the image, then turn the images into 10x10 grayscale images. Each image is then represented as a list of 100 numbers. The pixels are converted into numbers: -1 for white and 1 for black. The output is a list of 62 numbers where each number represents a different character. The number corresponding to the correct character is 1 and the rest are 0. After training, the highest output corresponds to the letter predicted by the network.

@@

First, we tested our implementation with simple functions: given a number, return the number, its negative, and half of the number. After we ironed out the implementation, the neural net was extremely accurate in approximating these functions. We could define a neural network with our own parameters as a function to be approximated, and the training network would exactly replicate the given network within 1\% of each parameter given enough cycles through the data. (In our case, 1000 cycles of 6 training examples through a network with 2 input, 2 hidden, and 2 output layers.)

\begin{wraptable}{l}{.45\textwidth}
\csvautotabular{pr_lows.csv}
\caption{Table 1 continued}
\vspace{-40px}
\end{wraptable}

@@

We found that neural networks are extremely efficient at sorting through large amounts of complicated data and finding the underlying function. However, they require a lot of data to be effective. In our case, increasing resolution of the image, and therefore increasing inputs, exponentially increases the amount of training time. On the whole, our lasting impression of neural networks is that they are an amazing tool to model convoluted functions given that we have a lot of data about the function.

\section*{Conclusion}
\label{Conclusion}

The big picture we take away from this class is that by applying data to adjust parameters, machines can be trained to interpret and predict data in ways that humans are unable to. We were taken aback by the complicated behavior: prediction, optimization, and character recognition, produced by a few simple lines of code combined with a large amount of data. We found that as we wrote our code, it became increasingly irrelevant what the numbers and weights of the algorithm were, but how those weights interacted with each other to produce a result. Through our survey, we discovered that this field of applied statistics and computer science called machine learning is very different from the traditional "programming" experiences we have had before. Besides the specific algorithms, we improved our implementation skills and learned new ways to think about handling and using data.

One of the primary issues we ran into with our course was a lack of data. Due to the nature of high school courses, we were could only use free, easily available, data. This meant we did not have access to large amounts of data that would have been required to test our programs more extensively and obtain better results. The lack of data was especially an issue with Artificial Neural Nets and Hidden Markov Models. In the case of Hidden Markov Models we were not even able to obtain a corpus to test our models and relied on Wikipedia data.

In general we, the students, learned the material and programmed by ourselves over the week and met on the weekend to consolidate our knowledge. We had weekly check ins after school with Mr. Petty to explain our progress. The independence that we had during the course made the class difficult at times, but also allowed us freedom to choose how we did things, and what we wanted to do with them. All of our projects were self driven, making them more fulfilling and more interesting than those of a regular computer science course. We would highly recommend future classes with similarly free structure and topics.
